---
title: | 
       | Laplacian Renormalization 
       | Group (LRG) Diseasome Analysis
pagetitle: "html_booklet"
author: "Begona Bolos & Colin Mclean"
date: "`r Sys.Date()`"
output:
  html_document:
    css: css/cruk.css
    number_sections: true
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
  word_document:
    toc: yes
    toc_depth: '4'
---


```{r setup, include=FALSE}

## load libraries
library(dplyr)
library(tidyr)
library(broom)
library(igraph)
library(poweRlaw)
library(Matrix)
library(expm)
library(ggplot2)
library(GGally)
library(scales)
library(latex2exp)
library(Deriv)
library(here)
library(Rbeast)
library(pracma) ## for zeta function
library(parallel)
library(future)
library(doFuture)
library(pso)

## set directory names
data.dir = here::here("datasets")

knitr::opts_chunk$set(cache=TRUE, echo=TRUE, warning=FALSE)
                      #root.dir=root.dir,
                      #project.dir=project.dir)

## In order to knit a Word doc, using kableExtra library for generating tables,
## 1) clear cache,
## 2) make sure this option is set before loading kableExtra in the Setup.R script.
options(kableExtra.auto_format=TRUE)

## Global options for displaying tables via DT::datatable
#options(DT.options=list(class='cell-border stripe',scrollx=TRUE))

options(DT.fillcontainer=TRUE, 
        DT.options=list(class='cell-border stripe',scrolly="400px", scrollx=TRUE))

## run Setup.R to load libraries globally - by specifying parameter local=knit_global()
#source(, local=knitr::knit_global())
```


```{r get_document_type, warning=FALSE, echo=FALSE, cache=TRUE}
doc.type=NULL 
output <- rmarkdown:::parse_yaml_front_matter(readLines(knitr::current_input()))$output
if( is.list(output) ){
   doc.type=names(output)[1]
   } else {
     doc.type=output[1]
   }
```

```{r, add_header_logo, warning=FALSE, echo=FALSE, cache=FALSE}
css.file=NULL
if(doc.type=="html_document"){
  ## get leading css file used
  css.file = rmarkdown:::parse_yaml_front_matter(readLines(knitr::current_input()))$output$html_document$css[1]
}

if( css.file=="css/cruk.css" ){
  htmltools::img(src =    knitr::image_uri(here::here("css/images/CRUK_SCOTLAND_C_Pos_RGB.png")),
               alt = "CRUK Scotland Centre logo",
               style = "position:absolute; top:0; right:0; padding:10px; height:80px; width:360px;")
}

if( css.file=="css/nhs.css" ){
htmltools::img(src = knitr::image_uri(here::here("css/images/NHS_Lothian.jpg")),
               alt = "NHS Lothian logo",
               style = "position:absolute; top:0; right:0; padding:10px; height:90px; width:90px;")
}

```

<br>
<br>

# Introduction {-}

> Laplacian renomalisation group (LRG) is a dimensional reduction technique inspired from physics. 
> The specific heat of a physical system, ie network, can be used to measure the point at which
> the system transitions from high-dimensional space to a coarse-dimensional space whilst 
> preserving the underlying toplogy of the system.
>
> [1] Villegas, P., Gili, T., Caldarelli, G. et al. Laplacian renormalization group for heterogeneous networks. Nat. Phys. 19, 445–450 (2023). https://doi.org/10.1038/s41567-022-01866-8
>

<br>
<br>

```{r}

## load function required to perform LRG on a network
source(here("core_functions.R"))

```

<br>
<br>

# Diseasome Network


>
> Load \& Build network
>


```{r}

load(here("data/diseasome.rda"))

gg = diseasome

## get Largest Connected Component
gg = lcc(gg)

## get degree distribution
deg = as.vector(degree(gg))

## Node of nodes
Nv = length(V(gg))

## set node names if null
if( is.null(V(gg)$name) ){ V(gg)$name = seq(1,Nv,1) }

L = get.L(gg)
E = get.eigen(L, only.values=FALSE, inv.vec=TRUE)

Emean = E[[1]]
Evec  = E[[2]]
Evinv = E[[3]]
```

<br>
<br>


>
> Eigen Density Plot
>


```{r}

pt1 = lambda_plot(df=eigen_density(Emean))

```


```{r}
plot(pt1)
```

```{r}
#library(dplyr)
# Create the first tibble with time intervals from 10^-4 to 10^4 with step 0.1
df1 <- tibble(tau = seq(10^-4, 10^4, by = 0.1))

# Create the second tibble with time intervals from 10^-4 to 1 with step 10^-2
df2 <- tibble(tau = seq(10^-4, 1, by = 10^-2))

# Combine the two tibbles
df  <- bind_rows(df1, df2)

# Remove duplicates and reorder the rows by tau in ascending order
df  <- df %>% 
       distinct() %>%  # Remove duplicates
       arrange(tau)    # Reorder by tau (ascending)
```

> Plot 1-S and find turning points

# Plot 1-S

```{r}

## Plot (1-S)
pt2 = ggplot(df, aes(x=tau))+
     geom_function(fun=S_wrapper, args=list(n=Nv,e=Emean), linewidth=1,
                  aes(color="1-S"))+
    labs(x = "log(tau)", y = "1-S")+
    scale_color_manual(values = c("1-S" = "blue"),
                       breaks = c("1-S"),
                       name = "")+ 
    scale_y_continuous(
    
    ## First axis
    name = "1-S",
    
    ) + 
    scale_x_log10(
        breaks = scales::trans_breaks("log10", function(x) 10^x),
        labels = scales::trans_format("log10", scales::math_format(10^.x))
    ) +
    annotation_logticks(sides="b", outside=TRUE)+
    coord_cartesian(clip = "off")+
    theme_light()+
    theme(legend.position = "right")
```

```{r}
pt2
```

# Plot C

> (1-S) and C curves together

```{r}

## create data for plot
y1      = S_wrapper(e=Emean, t=df$tau, n=Nv)
y2      = dS_dlogt_wrapper(e=Emean, t=df$tau, n=Nv)
not.nan = !is.nan(y2)
y1      = y1[not.nan]
y2      = y2[not.nan]
t.set   = df$tau[not.nan]

## build dataset
df_c    = data.frame(t=t.set, s=y1, c=y2)

```

```{r}

## prepare plot
tmp.y2 = y2

#cut.off=0.1#0.05
cut.off=0.001 #y2[1]
tmp.y2[tmp.y2<=cut.off]=cut.off

b2        = range(tmp.y2)
b3        = log(seq(b2[1],b2[2], length=4)) 
b3        = exp(b3)
b3        = signif(b3,2)
y2.breaks = unique(c(0,b3))
log.y2    = log(tmp.y2)

a         = range(y1)
b         = range(log.y2)
dat       = data.frame(x=t.set, y1=y1, y2=log.y2)

```

```{r}

## create plot
pt3 <- ggplot(dat, aes(x))+
     geom_line(aes(y=y1, color="1-S"), linewidth=1)+
     geom_line(aes(y=(y2-b[1])/diff(b), color="C log(N)"),linewidth=1, alpha=.8)+    
     #geom_line(aes(y=y2, color="C log(N)"),linewidth=1, alpha=.8)+    
    scale_color_manual(values = c("1-S"   = "blue",
                                "C log(N)" = "red"),
                       breaks = c("1-S","C log(N)"),
                       name = "")+ 
     labs(x = "log(tau)")+
     scale_x_log10()+
     scale_y_continuous(
         name = "1-S",
         labels=~round(.,2),
         breaks=seq(0,1,.25),
         sec.axis=sec_axis(~.*b2[2],
                           name = "C log(N)",
                           #labels = scales::label_scientific(digits = 1),
                           breaks = y2.breaks,
                           labels = y2.breaks
                           ))+
    annotation_logticks(sides="b", outside=TRUE)+
    coord_cartesian(clip = "off")+
    theme_light()+
    theme(legend.position = "right")
  
```

```{r}
pt3
```

# Plot dC/dt

> Plot dC/dt (method 1)

```{r}

## Plot C
pt4 = ggplot(df, aes(x=tau))+
  geom_function(fun=dC_dt_test_wrapper, args=list(n=Nv,e=Emean),
                  linewidth=1, aes(color="dC/dt"))+
    geom_hline(yintercept=0, color="black", linetype="dashed", linewidth=1)+
    labs(x = "log(tau)", y = "")+
    scale_color_manual(values = c("dC/dt" = "red"),
                       breaks = c("dC/dt"),
                       name = "")+ 
    scale_y_continuous(
    name = "",##dC/dt",
    ) + 
     scale_x_log10(
        breaks = scales::trans_breaks("log10", function(x) 10^x),
        labels = scales::trans_format("log10", scales::math_format(10^.x))
    ) +
    annotation_logticks(sides="b", outside=TRUE)+
    coord_cartesian(clip = "off")+
    theme_light()+
    theme(legend.position = "right")

```

```{r}
pt4
```

# Stationary points

> Where does dC/dt cross zero

```{r}
tt.steps = 1e-3
tt      = tibble(tau=seq(1e-4,1e1, tt.steps))
dc      = dC_dt_test_wrapper(n=Nv, t=tt$tau, e=Emean)
dc.diff = diff(sign(dc))
dc.indx = which(abs(dc.diff)==2)
tau.x   = NA
if( length(dc.indx) > 0 ){ tau.x   = tt$tau[dc.indx] }
```

```{r}
print(tau.x)
```

# Turning points

> Plot turning points of dC/dt:

```{r}

tps = turing.points(pt4, print=0)

```

```{r}
plot(tps$pt)
```


# Powerlaw Fit to Eigenvalues

## Basic Limits

```{r}

## Limits Eigenvalues
tau_min = 1/Emean[1]
tau_max = tau.x[1]

## Approximate C0 from spectral Dimensions
## C0 = sd/2 ~ -log(N)/log(fiedler_eigenvalue)
sd = compute_spectral_dim(Emean)
C0 = sd/2

```

# Fits 

```{r}

pl = fit_con_pl(Emean)

points = tps$tp
points = points[points$tp.x_axis>=tau_min & points$tp.x_axis<=tau_max,]


lrg.fits = list()
for( i in 1:nrow(points) ){
  t = points$tp.x_axis[i]
  lrg.fits[[i]] = eigen_loss(t=t, gg=gg, L=L, 
                             e=Emean, v=Evec, vinv=Evinv, 
                             complex=FALSE,
                             method="eigen",
                             expm_method="Higham08.b",
                             loss_test="js_dist",
                             mle=0,
                             tol=1e-5,
                             order=1,
                             plots=1)
}

```

# (alpha+1) Heuristic

```{r}

## Need to investigate these heuristics further
taus <- rep(NA,2)

taus[1] = df_c[which.min(df_c$c <= (pl$alpha+1)/log(Nv)),1]
taus[2] = df_c[which.min(df_c$c <= (pl$alpha+1)/log(Nv)^2),1]

lrg.fits2 = list()

for( i in 1:length(taus) ){
lrg.fits2[[i]] = eigen_loss(t=taus[i], gg=gg, L=L, 
                            e=Emean, v=Evec, vinv=Evinv, 
                            complex=FALSE,
                            method="eigen",
                            expm_method="Higham08.b",
                            loss_test="js_dist",
                            mle=0,
                            tol=1e-5,
                            order=1,
                            plots=1)
}
```

>
> Estimate tau using C0 = alpha + 1
> 1. Fit powerlaw to Eigenvalues to calculate alpha 
> 2. Estimate tau from C0 = alpha + 1


```{r}

#c_cut = max(df_c$c[1],df_c$c[dim(df_c)[1]])
c_cut = 10^-(2.5)
t_cut = range(df_c$t[df_c$c>=c_cut])

df2 = data.frame(tau=df$tau[df$tau >= t_cut[1] & df$tau < t_cut[2]])

## Plot log(N)*C
pt5 = ggplot(df2, aes(x=tau))+
     geom_function(fun=dS_dlogt_wrapper, args=list(e=Emean, n=Nv), 
                   linewidth=1, aes(color="log(N) C"))+
    geom_vline(xintercept=tau_min, colour="red3")+
    geom_vline(xintercept=taus[[1]], colour="blue")+
    geom_vline(xintercept=taus[[2]], colour="blue")+
    geom_vline(xintercept=points$tp.x_axis[1], colour="grey")+
    geom_vline(xintercept=points$tp.x_axis[2], colour="grey")+
    geom_vline(xintercept=points$tp.x_axis[3], colour="grey")+
    geom_vline(xintercept=tau_max, colour="red3")+
    geom_hline(yintercept=C0, color="orange2")+
    labs(x = "log(tau)", y = "log(N) C")+
    scale_color_manual(values = c("log(N) C" = "red"),
                       breaks = c("log(N) C"),
                       name = "")+ 
   scale_y_log10(
        expand = c(0, .1),
        breaks = scales::trans_breaks("log10", function(x) 10^x),
        labels = scales::trans_format("log10", scales::math_format(10^.x))
    ) +
    scale_x_log10(
        expand = c(0, 0),
        breaks = scales::trans_breaks("log10", function(x) 10^x),
        labels = scales::trans_format("log10", scales::math_format(10^.x))
    ) +
    annotation_logticks(sides="b", outside=TRUE)+
    coord_cartesian(clip = "off")+
    theme_light()+
    theme(legend.position = "right")

```

```{r}
plot(pt5)
```

```{r}
pt6 <- pt4+
  geom_vline(xintercept=tau_min, colour="red3")+
    geom_vline(xintercept=taus[[1]], colour="blue")+
    geom_vline(xintercept=taus[[2]], colour="blue")+
    geom_vline(xintercept=points$tp.x_axis[1], colour="grey")+
    geom_vline(xintercept=points$tp.x_axis[2], colour="grey")+
    geom_vline(xintercept=points$tp.x_axis[3], colour="grey")+
    geom_vline(xintercept=tau_max, colour="red3")
  
```

```{r}
pt6
```

# Scan time

```{r}

## Faster Way Using Parallelisation
t.lower = tau_min
t.upper = tau_max

## Create a cluster
cl <- makeCluster(detectCores() - 1)

test_t    <- seq(t.lower, t.upper, length.out=100)
test_loss <- parallel_eigen_loss(t=test_t, gg=gg, L=L, 
                                 e=Emean, v=Evec, vinv=Evinv, 
                                 complex=FALSE,
                                 method="eigen",
                                 expm_method="Higham08.b",
                                 loss_test="js_dist",
                                 mle=0,
                                 tol=1e-5,
                                 order=1,
                                 plots=0,
                                 cl=cl)

lrg.scan0 = data.frame(t=test_t, metric=test_loss)

## Stop the cluster
stopCluster(cl)
```


# Scan time plot

```{r}
metric="jsd"
x = lrg.scan0$t #lrg.scan0$df$t
y = lrg.scan0$metric #lrg.scan0$df[,metric]
df.scan = data.frame(x=x, y=y)

pt7 <- ggplot(df.scan, aes(x,y)) +
  geom_line(size=.8) + 
  ggtitle("Jensson-Shannon Distance between ecdf of network, \n and cdf of fitted lrg network")+
  labs(x="time", y="JSD")+
   geom_vline(xintercept=tau_min, colour="red3")+
    geom_vline(xintercept=taus[[1]], colour="blue")+
    geom_vline(xintercept=taus[[2]], colour="blue")+
    geom_vline(xintercept=points$tp.x_axis[1], colour="grey")+
    geom_vline(xintercept=points$tp.x_axis[2], colour="grey")+
    geom_vline(xintercept=points$tp.x_axis[3], colour="grey")+
    geom_vline(xintercept=tau_max, colour="red3")+
  theme_light() 

```

```{r}
pt7
```

# Scan Best time

```{r}

## Test Annealing algorithm
indx.best = which.min(lrg.scan0$metric)# lrg.scan0$df$jsd)
tau.best  = lrg.scan0$t[indx.best] #lrg.scan0$df$t[indx.best]

## Annealing algorithm Best Solution
scan.best  =   eigen_loss(t=tau.best, gg=gg, L=L, 
                         e=Emean, v=Evec, vinv=Evinv, 
                         complex=FALSE,
                         method="eigen",
                         expm_method="Higham08.b",
                         loss_test="js_dist",
                         mle=0,
                         tol=1e-5,
                         order=1,
                         plots=1)

```

```{r}
scan.best$plots[[1]]
```

```{r}
scan.best$plots[[2]]
```

```{r}
scan.best$plots[[3]]
```

# Annealing Algorithm

```{r}

## Define initial search range
test.tau = points[points$tp.x_axis<=tau.best,]
test_min = test.tau$tp.min.x_axis
test_max = .83 #test.tau$tp.max.x_axis

## Create a cluster
cl <- makeCluster(detectCores() - 1)

## Perform iterative optimization
lrg.ann <- iterative_eigen_optimization(
  t_lower = test_min, 
  t_upper = test_max, 
  gg = gg, 
  L = L, 
  e = Emean, 
  v = Evec, 
  vinv = Evinv, 
  complex=FALSE,
  method="eigen",
  expm_method="Higham08.b",
  loss_test="js_dist",
  mle=0,
  tol=1e-5,
  order=1,
  plots=0,
  max_iter = 20, 
  n_points = 10, 
  cooling_rate=0.85,
  dt=1,
  refine_factor = 0.5,
  cl=cl
)

## Print the result
cat("Optimized t:", lrg.ann$best_t, "with loss:", lrg.ann$best_loss, "\n")

## Stop the cluster
stopCluster(cl)

```


# Annealing Best Solution 

```{r}

## Annealing algorithm Best Solution
ann.best  =   eigen_loss(t=lrg.ann$best_t, gg=gg, L=L, 
                         e=Emean, v=Evec, vinv=Evinv, 
                         complex=FALSE,
                         method="eigen",
                         expm_method="Higham08.b",
                         loss_test="js_dist",
                         mle=0,
                         tol=1e-5,
                         order=1,
                         plots=1)

```

```{r}
ann.best$plots[[1]]
```

```{r}
ann.best$plots[[2]]
```

```{r}
ann.best$plots[[3]]
```

# Scan plot w. Annealing time

```{r}
metric="jsd"
x = lrg.scan0$t
y = lrg.scan0$metric
df.scan = data.frame(x=x, y=y)

pt8 <- ggplot(df.scan, aes(x,y)) +
  geom_line() + 
  ggtitle("Jensson-Shannon Distance between ecdf of network, \n and cdf of fitted lrg network")+
  labs(x="time", y="JSD")+
  geom_point(data=lrg.ann$df, aes(x=t_new,y=loss_new), color="red")+
  geom_line(data=lrg.ann$df, aes(x=t_new,y=loss_new), color="red")+
  theme_light() 

##plot(df.scan, type="l", xlab("time"), ylab(metric))
```

```{r}
pt8
```

# Particle Swarm Optimisation

```{r}

run_pso = 0

if( run_pso ){

## TEST
t.upper = test.tau$tp.min.x_axis
t.lower = .83

## Create a cluster
cl <- makeCluster(1) #detectCores() - 1)

## Perform PSO optimization using the parallelized loss function
result <- pso::psoptim(
  par = NA, ##initial guess
  fn = function(par) parallel_eigen_loss(t=par, gg=gg, L=L, 
                                 e=Emean, v=Evec, vinv=Evinv, 
                                 complex=FALSE,
                                 method="eigen",
                                 expm_method="Higham08.b",
                                 loss_test="js_dist",
                                 mle=0,
                                 tol=1e-5,
                                 order=1,
                                 plots=0,
                                 cl=cl),
  lower = t.lower,  # Lower bounds for t
  upper = t.upper,  # Upper bounds for t
  control = list(maxit = 100, trace = 1, abstol = 1e-5, vectorize=TRUE,
                 REPORT=1, trace.stats=TRUE, maxit.stagnate=5)  # PSO control options
)


## Print the result
cat("Optimized t:", result$par, "with loss:", result$value, "\n")

# Stop the cluster
stopCluster(cl)

}

```

# Best PSO time

```{r}

if( run_pso){
  
## Annealing algorithm Best Solution
pso.best  =   eigen_loss(t=result$par, gg=gg, L=L, 
                         e=Emean, v=Evec, vinv=Evinv, 
                         complex=FALSE,
                         method="eigen",
                         expm_method="Higham08.b",
                         loss_test="js_dist",
                         mle=0,
                         tol=1e-5,
                         order=1,
                         plots=1)

  
}

```

```{r}
if(run_pso){pso.best$plots[[1]]}
```

```{r}
if(run_pso){pso.best$plots[[2]]}
```

```{r}
if(run_pso){pso.best$plots[[3]]}
```

# Best LRG time on specific heat plot

```{r}

## create plot
pt9 <- ggplot(dat, aes(x))+
     geom_line(aes(y=y1, color="1-S"), linewidth=1)+
     geom_line(aes(y=(y2-b[1])/diff(b), color="C log(N)"),linewidth=1, alpha=.8)+    
     ##geom_vline(xintercept=lrg.ann$best_t, linetype="dashed", color="black", linewidth=1)+
    scale_color_manual(values = c("1-S"   = "blue",
                                "C log(N)" = "red"),
                       breaks = c("1-S","C log(N)"),
                       name = "")+ 
     labs(x = "log(tau)")+
     annotate("rect", xmin=tau_min, xmax=tau_max,
                     ymin=0, ymax=1, color="grey", alpha = .3)+
     geom_vline(xintercept=tau.best, linetype="dashed", color="grey30", linewidth=1)+
     geom_vline(xintercept=lrg.ann$best_t, linetype="dashed", color="black", linewidth=1)+
     scale_x_log10()+
     scale_y_continuous(
         name = "1-S",
         labels=~round(.,2),
         breaks=seq(0,1,.25),
         sec.axis=sec_axis(~.*b2[2],
                           name   = "C log(N)",
                          ##labels = scales::label_scientific(digits = 1),
                           breaks = y2.breaks,
                           labels = y2.breaks
                           ))+
    annotation_logticks(sides="b", outside=TRUE)+
    coord_cartesian(clip = "off")+
    theme_light()+
    theme(legend.position = "right")
  
```

```{r}
pt9
```


# Supernodes at best tau


```{r}

R = real.LRG(e=Emean, v=Evec, vinv=Evinv, L=L, t=lrg.ann$best_t, gg=gg, method="eigen", order=1)
table(R$supernodes)

```

# Plot Layer


```{r, fig.width=12}

## Expect one community after applying method one on pre-synaptic network
indx = which(R$supernodes>1)

names(indx) = seq(1,length(indx),1)

## Set node colors if highlighting the one community found
R$mapping %>% 
    data.frame %>%
    mutate(color = match(as.numeric(V2), indx)) %>%
    mutate(color = ifelse(is.na(color),0,color)) %>%
    select(color) -> node_color

## Assign colors
node_color$color_hex <- generate_colors(node_color$color)

## Update Vertex attribute for plotting
gg <- set_vertex_attr(gg, "cluster", V(gg), as.vector(node_color$color_hex))

# Build plot
pt9 <- ggnet2(gg, edge.color = "grey80", node.color = "cluster",
              palette = unique(node_color$color_hex), size = 1)


```

```{r}
pt9
```


