---
title: | 
       | Laplacian Renormalization 
       | Group (LRG) Presynaptic PPI Analysis
pagetitle: "html_booklet"
author: "Begona Bolos & Colin Mclean"
date: "`r Sys.Date()`"
output:
  html_document:
    css: css/cruk.css
    number_sections: true
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
  word_document:
    toc: yes
    toc_depth: '4'
---


```{r setup, include=FALSE}

## load libraries
library(dplyr)
library(tidyr)
library(broom)
library(igraph)
library(poweRlaw)
library(Matrix)
library(expm)
library(ggplot2)
library(GGally)
library(latex2exp)
library(Deriv)
library(here)
library(Rbeast)
library(pracma) ## for zeta function
library(parallel)
library(pso)

## set directory names
data.dir = here::here("datasets")

knitr::opts_chunk$set(cache=TRUE, echo=TRUE, warning=FALSE)
                      #root.dir=root.dir,
                      #project.dir=project.dir)

## In order to knit a Word doc, using kableExtra library for generating tables,
## 1) clear cache,
## 2) make sure this option is set before loading kableExtra in the Setup.R script.
options(kableExtra.auto_format=TRUE)

## Global options for displaying tables via DT::datatable
#options(DT.options=list(class='cell-border stripe',scrollx=TRUE))

options(DT.fillcontainer=TRUE, 
        DT.options=list(class='cell-border stripe',scrolly="400px", scrollx=TRUE))

## run Setup.R to load libraries globally - by specifying parameter local=knit_global()
#source(, local=knitr::knit_global())
```


```{r get_document_type, warning=FALSE, echo=FALSE, cache=TRUE}
doc.type=NULL 
output <- rmarkdown:::parse_yaml_front_matter(readLines(knitr::current_input()))$output
if( is.list(output) ){
   doc.type=names(output)[1]
   } else {
     doc.type=output[1]
   }
```

```{r, add_header_logo, warning=FALSE, echo=FALSE, cache=FALSE}
css.file=NULL
if(doc.type=="html_document"){
  ## get leading css file used
  css.file = rmarkdown:::parse_yaml_front_matter(readLines(knitr::current_input()))$output$html_document$css[1]
}

if( css.file=="css/cruk.css" ){
  htmltools::img(src =    knitr::image_uri(here::here("css/images/CRUK_SCOTLAND_C_Pos_RGB.png")),
               alt = "CRUK Scotland Centre logo",
               style = "position:absolute; top:0; right:0; padding:10px; height:80px; width:360px;")
}

if( css.file=="css/nhs.css" ){
htmltools::img(src = knitr::image_uri(here::here("css/images/NHS_Lothian.jpg")),
               alt = "NHS Lothian logo",
               style = "position:absolute; top:0; right:0; padding:10px; height:90px; width:90px;")
}

```

<br>
<br>

# Introduction {-}

> Laplacian renomalisation group (LRG) is a dimensional reduction technique inspired from physics. 
> The specific heat of a physical system, ie network, can be used to measure the point at which
> the system transitions from high-dimensional space to a coarse-dimensional space whilst 
> preserving the underlying toplogy of the system.
>
> [1] Villegas, P., Gili, T., Caldarelli, G. et al. Laplacian renormalization group for heterogeneous networks. Nat. Phys. 19, 445–450 (2023). https://doi.org/10.1038/s41567-022-01866-8
>

<br>
<br>

```{r}

## load function required to perform LRG on a network
source(here("core_functions.R"))

```

<br>
<br>

# Presynaptic PPI Network

>
> Load \& Build network
>

```{r}

gg = read_graph(here("datasets/PPI_Presynaptic.gml"), format="gml") 

## get Largest Connected Component
gg = lcc(gg)

## Node of nodes
Nv = length(V(gg))

## set node names if null
if( is.null(V(gg)$name) ){ V(gg)$name = seq(1,Nv,1) }

L = get.L(gg)
E = get.eigen(L, only.values=FALSE, inv.vec=TRUE)

Emean = E[[1]]
Evec  = E[[2]]
Evinv = E[[3]]
```

<br>
<br>


>
> Plots
>

>
> Eigenvalues against degree
>

```{r}
#library(dplyr)
# Create the first tibble with time intervals from 10^-4 to 10^4 with step 0.1
df1 <- tibble(tau = seq(10^-4, 10^4, by = 0.1))

# Create the second tibble with time intervals from 10^-4 to 1 with step 10^-2
df2 <- tibble(tau = seq(10^-4, 1, by = 10^-2))

# Combine the two tibbles
df  <- bind_rows(df1, df2)

# Remove duplicates and reorder the rows by tau in ascending order
df  <- df %>% 
       distinct() %>%  # Remove duplicates
       arrange(tau)    # Reorder by tau (ascending)
```

> Compute S and C

```{r}

## Possible Specific Heat Capacity Calculations
S  = S_wrapper(e=Emean, t=df$tau, n=Nv, negate=0)
C1 =-diff(S)/diff(log(df$tau))
C1 = c(C1,C1[length(C1)])

C2 = sapply(df$tau, function(i) -dS_dlogt(e=Emean, t=i, n=Nv))

##plot(x=(df$tau), y=C1, type="l", ylim=c(0,1), xlim=c(0,10), col="blue")+lines(x=(df$tau), y=C2, col="red")

pl = eigen.fit2(Emean)

## plot
lambda_fit_plot(pl$df, xmin=pl$xmin, alpha=pl$alpha)

#P = log(pl$df)

#plot(P[,1], P[,2], type = "p", col = "blue", pch = 16,
#     xlab = "log10(Laplacian Eigenvalues)", ylab = "log10(CDF Probabilities)",
#     main = "Log-Log Plot with Power-Law Fits")

# Step 5: Incrementally fit power laws
for (i in seq(10, 600, by = 10)) {
  # Subset data
  x_subset <- P[1:i,1]
  y_subset <- P[1:i,2]
  
  # Fit a linear model (power-law on log-log scale)
  model <- lm(y_subset ~ x_subset)
  
  # Extract coefficients
  a <- coef(model)[1]  # Intercept
  b <- coef(model)[2]  # Slope
  
  # Generate fitted line
  fitted_y <- a + b * P[,1]#x_subset
  
  # Plot the fitted line
  lines(P[,1], fitted_y, col = "red", lty = 2)
}

# Step 5: Finite-Size Scaling Analysis
lambda_F  <- sort(Emean)[2]
alpha     <- 2
tau_F     <- alpha/lambda_F
ds_values <- seq(1, 20, 0.1)  # Test various d_S
C_values  <- sapply(ds_values, function(ds) C_tau(tau_F, ds))

# Plot C(τ_F) as a function of d_S
plot(ds_values, C_values, type = "l", col = "blue", lwd = 2,
     xlab = "d_S (Spectral Dimension)", ylab = "C(τ_F)",
     main = "C(τ_F) vs. Spectral Dimension")


```

```{r}
## Set parameters
params = list()
#params[[1]]=Emean
#params[[2]]=Evec 
#params[[3]]=Evinv 
params[[1]]=FALSE #complex=FALSE 
params[[2]]="eigen" #method="eigen"
params[[3]]="Higham08.b" #expm_method="Higham08.b" 
params[[4]]="ks_dist" #js_dist"#"kl_div"
params[[5]]=1 #mle=1
params[[6]]=1e-5 #tol=1e-5 
params[[7]]=1 #order=1 
params[[8]]=0 #plots=0
names(params) = c("complex", "method","expm_method", "loss_test", "mle", "tol", "order", "plots")
```

```{r}

## Faster Way Using Parallelisation
t.lower = 1/Emean[1] #0.001
t.upper = 1

## Create a cluster
cl <- makeCluster(detectCores() - 1)

test_t    <- seq(t.lower, t.upper, length.out=100)
test_loss <- parallel_eigen_loss(t=test_t, gg=gg, L=L, 
                                 e=Emean, v=Evec, vinv=Evinv, 
                                 complex=params[[1]], 
                                 method=params[[2]],
                                 expm_method=params[[3]],
                                 loss_test=params[[4]],
                                 mle=params[[5]],
                                 tol=params[[6]],
                                 order=params[[7]],
                                 plots=params[[8]],
                                 cl=cl)

lrg.scan0 = data.frame(t=test_t, metric=test_loss)

## Stop the cluster
stopCluster(cl)
```

 
```{r}

## Create a cluster
cl <- makeCluster(detectCores() - 1)

## Define initial search range
t_lower <- 1/Emean[1] #0.01
t_upper <- 1#8 #tau.x #1

## Perform iterative optimization
lrg.ann <- iterative_eigen_optimization(
  t_lower = t_lower, 
  t_upper = t_upper, 
  gg = gg, 
  L = L, 
  e = Emean, 
  v = Evec, 
  vinv = Evinv, 
  complex=params[[1]], 
  method=params[[2]],
  expm_method=params[[3]],
  loss_test=params[[4]],
  mle=params[[5]],
  tol=params[[6]],
  order=params[[7]],
  plots=params[[8]],
  max_iter = 20, 
  n_points = 10, 
  cooling_rate=0.85,
  dt=1,
  refine_factor = 0.5,
  cl=cl
)

## Print the result
cat("Optimized t:", lrg.ann$best_t, "with loss:", lrg.ann$best_loss, "\n")

## Stop the cluster
stopCluster(cl)

```

```{r}
metric="jsd"
x = lrg.scan0$t 
y = lrg.scan0$metric 
df.scan = data.frame(x=x, y=y) %>%
          filter( !is.infinite(y) | is.na(y) )

df.ann = lrg.ann$df %>% select(t_new, loss_new) %>% 
         filter(!is.infinite(t_new))

pt5 <- ggplot(df.scan, aes(x,y)) +
  geom_line() + 
  ggtitle("Jensson-Shannon Distance between ecdf of network, \n and cdf of fitted lrg network (eigenvalues)")+
  labs(x="time", y="JSD")+
  geom_point(data=df.ann, aes(x=t_new,y=loss_new), color="red")+
  geom_line(data=df.ann, aes(x=t_new,y=loss_new), color="red")+
  theme_light() 

```

```{r}
plot(pt5)
```

```{r}

best=eigen_loss(t=lrg.ann$best_t, 
                gg=gg, L=L, e=Emean, 
                v=Evec, vinv=Evinv, 
                complex=params[[1]], 
                method=params[[2]],
                expm_method=params[[3]],
                loss_test=params[[4]],
                mle=params[[5]],
                tol=params[[6]],
                order=params[[7]],
                plots=1)
```

```{r}
plot(best$plots[[1]])
```

```{r}
plot(best$plots[[2]])
```

```{r}

Ct  = sapply(df$tau, function(i) -dS_dlogt(e=Emean, t=i, n=Nv))

dat = data.frame(x=df$tau, y=Ct, y1=Ct*log(Nv), y2=log(Ct*log(Nv)))

```

```{r}

dat = dat[!is.infinite(dat$y),]
dat = dat[!is.infinite(dat$y1),]
dat = dat[!is.infinite(dat$y2),]

## create plot
pt3 <- ggplot(dat, aes(x,y1))+
     geom_hline(yintercept=(best$params$spec_dim/2), color="orange2")+
     geom_hline(yintercept=(best$params$C0), color="orange")+
     geom_vline(xintercept=lrg.ann$best_t)+
     geom_line(aes(color="C log(N)"),linewidth=1, alpha=.8)+    
     scale_color_manual(values = c("C log(N)" = "red"),
                       breaks = c("C log(N)"),
                       name = "")+ 
     labs(x = "log(tau)")+
     scale_x_log10()+
     scale_y_continuous(
         name   = "C log(N)",
         #labels = ~round(.,2),
         #breaks = seq(-10,2,.5),
         labels = scales::label_scientific(digits = 2))+
    #annotation_logticks(sides="b", outside=TRUE)+
    coord_cartesian(clip = "off")+
    theme_light()+
    theme(legend.position = "right")
  
```

```{r}
plot(pt3)
```

```{r}

## Plot dC/dt
pt4 = ggplot(df, aes(x=tau))+
  geom_function(fun=dC_dt_test_wrapper, args=list(n=Nv,e=Emean),
                  linewidth=1, aes(color="dC/dt"))+
    geom_hline(yintercept=0, color="black", linetype="dashed", linewidth=1)+
    geom_vline(xintercept=lrg.ann$best_t)+
    labs(x = "log(tau)", y = "")+
    scale_color_manual(values = c("dC/dt" = "red"),
                       breaks = c("dC/dt"),
                       name = "")+ 

    scale_y_continuous(
    name = "",##dC/dt",
    ) + 
     scale_x_log10(
        breaks = scales::trans_breaks("log10", function(x) 10^x),
        labels = scales::trans_format("log10", scales::math_format(10^.x))
    ) +
    annotation_logticks(sides="b", outside=TRUE)+
    coord_cartesian(clip = "off")+
    theme_light()+
    theme(legend.position = "right")
```

```{r}
pt4
```