---
title: | 
       | Early Onset CRC
       | PandaR Study
pagetitle: "html_booklet"
author: "Colin Mclean"
date: "`r Sys.Date()`"
output:
  html_document:
    css: css/cruk.css
    number_sections: true
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
  word_document:
    toc: yes
    toc_depth: '4'
---


```{r setup, include=FALSE}

## load libraries
library(NanoStringNCTools)
library(ggplot2)
library(ggridges)
library(here)
library(ggthemes)
library(ggiraph)
library(MASS)
library(stringr)
library(dplyr)
library(purrr)
library(matrixStats)
library(patchwork)
library(ggfortify)
library(gridExtra)
library(grid)
library(data.table)
library(tinytex)
library(latex2exp)
library(igraph)
library(Deriv)
library(readr)
library(tidyr)


## set directory names
data.dir = here::here("../datasets")

knitr::opts_chunk$set(cache=TRUE, echo=TRUE, warning=FALSE)
                      #root.dir=root.dir,
                      #project.dir=project.dir)

## In order to knit a Word doc, using kableExtra library for generating tables,
## 1) clear cache,
## 2) make sure this option is set before loading kableExtra in the Setup.R script.
options(kableExtra.auto_format=TRUE)

## Global options for displaying tables via DT::datatable
#options(DT.options=list(class='cell-border stripe',scrollx=TRUE))

options(DT.fillcontainer=TRUE, 
        DT.options=list(class='cell-border stripe',scrolly="400px", scrollx=TRUE))

## run Setup.R to load libraries globally - by specifying parameter local=knit_global()
#source(, local=knitr::knit_global())
```


```{r get_document_type, warning=FALSE, echo=FALSE, cache=TRUE}
doc.type=NULL 
output <- rmarkdown:::parse_yaml_front_matter(readLines(knitr::current_input()))$output
if( is.list(output) ){
   doc.type=names(output)[1]
   } else {
     doc.type=output[1]
   }
```

```{r, add_header_logo, warning=FALSE, echo=FALSE, cache=FALSE}
css.file=NULL
if(doc.type=="html_document"){
  ## get leading css file used
  css.file = rmarkdown:::parse_yaml_front_matter(readLines(knitr::current_input()))$output$html_document$css[1]
}

if( css.file=="css/cruk.css" ){
  htmltools::img(src =    knitr::image_uri(here::here("css/images/CRUK_SCOTLAND_C_Pos_RGB.png")),
               alt = "CRUK Scotland Centre logo",
               style = "position:absolute; top:0; right:0; padding:10px; height:80px; width:360px;")
}

if( css.file=="css/nhs.css" ){
htmltools::img(src = knitr::image_uri(here::here("css/images/NHS_Lothian.jpg")),
               alt = "NHS Lothian logo",
               style = "position:absolute; top:0; right:0; padding:10px; height:90px; width:90px;")
}

```


```{r}

## load function required to perform LRG on a network
source(here("nanostring_RUV_functions.R"))
source("core_functions.R")
source("panda_ppi.R")

```

# Load Test Data

```{r}

## Read in normalised expression data 
x=readRDS(here("pandaR_study.RDS"))

raw_data  = x$expr_dat
pat_anno  = x$pat_anno
gene_anno = x$gene_anno
ppi       = x$ppi_dat
tf_motif  = x$tf_dat
df        = x$df

```

# Network Reconstruction Approach

>
> 1. Use nSolver Normalised Data. 
>
> 2. Use Stage 3 and 4 patients to build aggregate / population 
>    networks.
>
> 3. Use PANDA regulatory network construction algorithm 
> 
> 4. The PANDA algorithm requires three input networks:
> 
>  4.1. Gene-Transcript network (Broad Institute's dataset). 
>       Need to ask Susanti which transcrition factors (TFs) from her gene set 
>       are expected to influence genes, and which TFs from her gene set are not.
>
>
>  4.2. PCC co-expression network for genes
> 
>  4.3. A PPI reference network (HumanNet-XN v2 network (https://www.inetbio.org/humannetv2/))
>
> Gene Sets
>
> 1. https://github.com/mahmoodlab/MCAT/blob/master/datasets_csv_sig/signatures.csv
>
> 2. https://www.xiejjlab.bio/MTFAP/browse
>
> 3. https://github.com/lawrenson-lab/CaCTS/tree/master


# PPI Network

```{r}

## Get set of unique gene entrez IDs
gn_ids = gene_anno %>%
         separate_rows(entrez, sep=";") %>%
         dplyr::select(entrez) %>%
         distinct() %>%
         filter(!is.na(entrez)) %>%
         unlist() %>%
         as.vector()

## Get set of unique gene names
gn_name     = gene_anno %>%
              separate_rows(gene_name, sep=";") %>%
              dplyr::select(gene_name) %>%
              distinct() %>%
              filter(!is.na(gene_name)) %>%
              unlist() %>%
              as.vector()

## Get set of unique gene entrez IDs
tf_ids = gene_anno %>%
         separate_rows(entrez, sep=";") %>%     
         filter(tf_inData) %>%
         dplyr::select(entrez) %>%
         filter(!is.na(entrez)) %>%
         distinct() %>%
         unlist() %>%
        as.vector()

## Get set of unique gene names
tf_name     = gene_anno %>%
              separate_rows(gene_name, sep=";") %>%
              filter(tf_inData) %>%
              dplyr::select(gene_name) %>%
              distinct() %>%
              filter(!is.na(gene_name)) %>%
              unlist() %>%
              as.vector()

```


```{r, eval=FALSE}

## Load PPIs
ppi   = read.csv(here("PPI/HumanNet-PI.tsv"), sep="\t", header=FALSE, skip=1)

```

```{r}

edges = ppi[(ppi$V1 %in% gn_ids & ppi$V2 %in% gn_ids),1:2]

edges[] <- lapply(edges, as.character)

edge_names <- cbind(apply(edges, 2, function(x) 
                                     gn_name[match(x,gn_ids)]))

gg    = igraph::graph_from_edgelist(as.matrix(edge_names), directed=FALSE)
gg2   = lcc(gg)

d      = sort(degree(gg2)) 
df_deg = data.frame(x=d,y=data_cdf_probs(d))

gene_anno$degree = df_deg[match(rownames(gene_anno), rownames(df_deg)),"x"]

pt15 <- lambda_plot(df_deg)

## Extract edges from 
crc_ppi = as_edgelist(gg2) %>%
          data.frame() %>%
          mutate(value = 1)

```

```{r}
plot(pt15)
```

# TF PPI Net

```{r}

edges = ppi[(ppi$V1 %in% tf_ids & ppi$V2 %in% tf_ids),1:2]

edges[]    <- lapply(edges, as.character)
edge_names <- cbind(apply(edges, 2, function(x) 
                                     tf_name[match(x,tf_ids)]))

gg.tf = igraph::graph_from_edgelist(as.matrix(edge_names), directed=FALSE)
gg3   = lcc(gg.tf)

d      = sort(degree(gg3)) 
df_deg = data.frame(x=d,y=data_cdf_probs(d))
 
gene_anno$TFdegree = df_deg[match(rownames(gene_anno), rownames(df_deg)),"x"]
 
pt16 <- lambda_plot(df_deg)

## Extract edges from 
tf_ppi = as_edgelist(gg3) %>%
          data.frame() %>%
          mutate(value = 1)

```

```{r}
plot(pt16)
```

# TF motif

```{r, eval=FALSE}

# ## Check number of unique detection methods
# unique_methods <- tf_motif %>%
#     separate_rows(Detection.method, sep = ";") %>%  # Split rows by ";"
#     distinct(Detection.method) %>%  # Remove duplicates
#     arrange(Detection.method)  # Sort alphabetically


# Filter TF_motif data based on most reliable methods
method_set = c("chromatin immunoprecipitation assay", "DNase I footprinting", "electrophoretic mobility shift assay")

tf_motif <- read.delim2(here("PPI/TFLink_Homo_sapiens_interactions_All_simpleFormat_v1.0.tsv"), 
                        sep="\t", header=TRUE,
                        stringsAsFactors=FALSE) %>%
    as_tibble() %>%
    mutate(n_pubs    = str_count(PubmedID, ";")) %>%
    mutate(n_targets = str_count(Name.Target, ";")) %>%
    separate_rows(Detection.method, sep = ";") %>%  # Split multiple methods into separate rows
    filter(Detection.method %in% method_set)  %>% # Keep only rows that match method_set
    filter(n_pubs > 1) %>%
    separate_rows(Name.Target, sep = ";") #%>%  # Split targets into separate rows

```

```{r}

## Filter for TFs in signature dataset
tf_motif_sub  <- tf_motif %>%
                 filter(Name.TF %in% tf_name) %>%
                 filter(Name.Target %in% gn_name)


tf_priors = tf_motif_sub %>% 
            group_by(Name.TF, Name.Target) %>%
            mutate(dup_count = n()) %>% 
            filter(dup_count==1) %>%
            ungroup() %>%
            dplyr::select(Name.TF, Name.Target) %>%
            mutate(value = 1) %>%
            data.frame()


tf_ppi_sub = tf_ppi[tf_ppi[,2] %in% union(tf_ppi[,1], tf_priors[,1]),]

```

# pandaR

```{r}

fData           = raw_data@featureData@data
expr_genes_indx = !fData$CodeClass %in% c("Positive", "Negative")
expr_genes      = fData$GeneName[expr_genes_indx]

# Function to split gene names correctly
split_gene_names <- function(gene) {
  if(grepl("/", gene)){
    parts <- unlist(strsplit(gene, "/"))  # Split at '/'
    ## First gene is the original first part
    gene1   <- parts[1]
    postfix <- parts[2]
    N <- str_length(gene1)
    n <- str_length(postfix)
    ## Second gene keeps the prefix but replaces the number
    gene2 <- paste0(str_sub(gene1, 1, (N-n)), postfix)  # Reconstruct second gene

  return(paste0(c(gene1, gene2), collapse=";"))
  }
  return(gene)
}

```

```{r}

## Normalise using nSolver method
raw_data <- NanoStringNCTools::normalize(raw_data, type="nSolver", fromELT = "exprs", toELT = "exprs_norm")

## Extract log normalised data
x        <- Biobase::assayDataElement(raw_data, elt = "exprs_norm")

## Rename Samples and 
colnames(x)=df$id
rownames(x)=fData$GeneName

## remove positive/negative controls
x = x[expr_genes_indx,]

## How to deal with isoform's like "CCL3/L1". Approach here is 
## is to duplicate the isoform expression row to create two rows for 
## "CCL3" and "CCL1"
x <- x %>%
     data.frame() %>%
     tibble::rownames_to_column("Gene") %>%   # Convert row names to a column
     dplyr::mutate(Gene = sapply(Gene, function(x) split_gene_names(x)[[1]])) %>%
     tidyr::separate_rows(Gene, sep = ";") %>%  # Split and expand rows
     tibble::column_to_rownames("Gene")  # Restore row names

group  = c("early", "late")
strata = c("severe"=list(c(3,4)), "moderate"=list(c(1,2)))

cohorts = list() 

for( i in group ){
  for( j in 1:length(strata) ){
    cohort_name = sprintf("%s_%s", i, names(strata)[j])
    cohorts[[cohort_name]] = data.frame(x[,(grepl(i,df$group) & df$stage >= strata[[j]])])
  }
  
}

```


```{r}

gene.reg.net = list()

## run pandas
for( i in 1:length(cohorts)){
  gene.reg.net[[i]]=pandaR::panda(motif=tf_priors, 
                                  expr=cohorts[[i]], 
                                  ppi=tf_ppi_sub)#, progress=TRUE)
  names(gene.reg.net)[i] = names(cohorts)[i]
}

```


```{r}

dims = dim(gene.reg.net[[1]]@regNet)
df1  = matrix(0, dims[1], dims[2])
rownames(df1) = rownames(gene.reg.net[[1]]@regNet)

for( i in 1:dims[1] ){
  df1[i,] = gene.reg.net[[1]]@regNet[i,] - gene.reg.net[[3]]@regNet[i,]
}

pt17 <- df1 %>%
        reshape2::melt(.) %>%
        arrange(Var1) %>%
        ggplot(., aes(x=value, y=Var1, fill=Var1))+
        geom_density_ridges(scale = 1.2, alpha = 0.7)+
        theme_minimal()+
        xlim(c(-0.5,0.5))+
        labs(title = "",
        x = "ealry(extreme) - late(extreme)",
        y = "Transcription Factor") +
        theme(legend.position = "none") # Remove legend for clarity


dims = dim(gene.reg.net[[1]]@regNet)
df2  = matrix(0, dims[1], dims[2])
rownames(df2) = rownames(gene.reg.net[[1]]@regNet)

for( i in 1:dims[1] ){
  df2[i,] = gene.reg.net[[2]]@regNet[i,] - gene.reg.net[[4]]@regNet[i,]
}

pt18 <- df2 %>%
        reshape2::melt(.) %>%
        arrange(Var1) %>%
        ggplot(., aes(x=value, y=Var1, fill=Var1))+
        geom_density_ridges(scale = 1.2, alpha = 0.7)+
        theme_minimal()+
        xlim(c(-0.5,0.5))+
        labs(title = "",
        x = "ealry(moderate) - late(moderate)",
        y = "Transcription Factor") +
        theme(legend.position = "none") # Remove legend for clarity


```

```{r, fig.width=10, fig.height=10}
## Use patchwork package to display all plots 
(pt17 | pt18) 
```

```{r}

dims = dim(gene.reg.net[[1]]@regNet)
df   = matrix(0, dims[1], dims[2])
rownames(df) = rownames(gene.reg.net[[1]]@regNet)

for( i in 1:dims[1] ){
  df[i,] = gene.reg.net[[1]]@regNet[i,] - gene.reg.net[[2]]@regNet[i,]
}

pt19 <- df %>%
        reshape2::melt(.) %>%
        arrange(Var1) %>%
        ggplot(., aes(x=value, y=Var1, fill=Var1))+
        geom_density_ridges(scale = 1.2, alpha = 0.7)+
        theme_minimal()+
        xlim(c(-0.5,0.5))+
        labs(title = "",
        x = "ealry(extreme) - early(moderate)",
        y = "Transcription Factor") +
        theme(legend.position = "none") # Remove legend for clarity


dims = dim(gene.reg.net[[1]]@regNet)
df2  = matrix(0, dims[1], dims[2])
rownames(df2) = rownames(gene.reg.net[[1]]@regNet)

for( i in 1:dims[1] ){
  df2[i,] = gene.reg.net[[3]]@regNet[i,] - gene.reg.net[[4]]@regNet[i,]
}

pt20 <- df2 %>%
        reshape2::melt(.) %>%
        arrange(Var1) %>%
        ggplot(., aes(x=value, y=Var1, fill=Var1))+
        geom_density_ridges(scale = 1.2, alpha = 0.7)+
        theme_minimal()+
        xlim(c(-0.5,0.5))+
        labs(title = "",
        x = "late(severe) - late(moderate)",
        y = "Transcription Factor") +
        theme(legend.position = "none") # Remove legend for clarity


```

```{r, fig.width=10, fig.height=10}
## Use patchwork package to display all plots 
(pt19 | pt20) 
```

# Panda for PPI

>
> Study to use panda algorithm to iteratively predict an early and late onset 
> ppi network given expression data, tf-gene interaction data   
>

```{r}

## Network for Early Onset Extreme using Panda
eo_extreme = panda_ppi(motif=tf_priors, 
                       expr=cohorts[[1]], 
                       ppi=crc_ppi)

eo_exteme_net = eo_extreme$PPINet %>%
                melt() %>%
                filter(value>0.5)

gg_eo = graph_from_data_frame(eo_exteme_net) %>%
        lcc()

d      = sort(degree(gg_eo)) 
df_deg = data.frame(x=d,y=data_cdf_probs(d))

pt1 <- lambda_plot(df_deg)



## Network for Late Onset Extreme using Panda
lo_extreme = panda_ppi(motif=tf_priors, 
                       expr=cohorts[[3]], 
                       ppi=crc_ppi)

lo_exteme_net = lo_extreme$PPINet %>%
                melt() %>%
                filter(value>0.5)


gg_lo = graph_from_data_frame(lo_exteme_net) %>%
        lcc()

d      = sort(degree(gg_lo)) 
df_deg = data.frame(x=d,y=data_cdf_probs(d))

pt2 <- lambda_plot(df_deg)

```

# Early Onset Network

```{r}
summary(gg_eo)
```

```{r}
pt1
```

# Late Onset Network

```{r}
summary(gg_lo)
```

```{r}
pt2
```