---
title: "BioNAR parallel execution"
author: "Colin Mclean, Anatoly Sorokin, T. I. Simpson, J. Douglas Armstrong, Oksana Sorokina"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BioNAR parallel execution}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
library(knitr)
library(BioNAR)
library(synaptome.db)
library(ggplot2)
library(pander)
library(ggrepel)
library(randomcoloR)
```

# Introduction
There are several functions in the BioNAR package that takes substantial amount 
of time either because of complex calculations or because of repetitive random
perturbation of the graph under investigation. The later functions could benefit
a lot from the parallelism, taking into account that most contemporary computer
have more than one core. BioNAR support multicore execution since version 1.3.7,
via 
[BiocParallel](https://bioconductor.org/packages/release/bioc/html/BiocParallel.html).
This vignette describe which functions could be run within parallel execution
environment and how to do it most efficiently.

# Linux and MacOS
Linux and MacOS support `MulticoreParam`, which is the most efficient way to run
randomisation jobs in parallel:
>"MulticoreParam uses ‘forked’ processes with ‘copy-on-change’ semantics – memory is only copied when it is changed. This makes it very efficient to invoke compared to other back-ends."
([via](https://bioconductor.org/packages/release/bioc/vignettes/BiocParallel/inst/doc/Introduction_To_BiocParallel.html))

# Windows
Windows do not supports 'forked' processes, so the only parallel environment
available is `SnowParam`, which emulate cluster of workstations by setting up 
several instances and run code there. That approach implies huge overhead, which
could be several orders of magnitude higher than the execution time, so by
default on Windows BioNAR uses `SerialParam` environment and execute all code
on one core. In the following we will discuss when the `SnowParam` is useful despite the overhead and how to set up parallel environment in a way that maximize benefits
from parallel execution.




Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
