---
title: "BioNAR: Biological Network Analysis in R"
author: "Colin Mclean, Anatoly Sorokin, T. I. Simpson, J. Douglas Armstrong, Oksana Sorokina"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BioNAR: Biological Network Analysis in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
library(knitr)
library(BioNAR)
library(synaptome.db)
library(ggplot2)
library(pander)
library(ggrepel)
library(randomcoloR)
```

# Introduction

Most of proteomic studies usually result with a massive list of proteins being
identified within a specific tissue, compartment or cell type, occasionally 
accompanied by some qualitative or quantitative information. To obtain m
eaningful biological insight from this data, one usually needs to process the 
data in several stages to identify possible structural and/or functional dependencies.

One of the most popular ways of analysing the proteomic data is the building 
of the protein-protein interaction networks, to study how its topology 
influences functional annotation mapped onto it. 

Many existing packages support different steps of the network building and 
analysis process, but there are few packages that combine network building 
and analysis methodology into a single pipeline. 

We designed the BioNAR to combine the necessary network building and analysis
functionality, complementing existing R packages and filling the methodological
gaps necessary to interrogate biomedical networks with respect to functional 
and disease domains. For that purpose, we didn’t not implement the network
reconstruction (unless it is synaptic network), given that other tools like
Cytoscape and Network Analyst can do this already. Instead, we implemented a
detailed topologically-based network analysis package, enabling the researcher 
to load networks generated from the lab’s own meta-data, thus making the tool 
as widely applicable and flexible as possible.

# Overview of capabilities
The package’s pipeline starts with loading/building the graph of interest and
annotating its vertices with available metadata [annotate_vertex]. 

It follows with the analysis of general network properties, such as estimating 
a network’s “scale-free” property. For this we used the R “PoweRlaw” package
(version 0.50.0) (Gillespie, 2015) [FitDegree] and a network entropy analysis
(Teschendorff et al, 2014) [getEntopy].

Package allows estimating of the main network vertex centrality measures, 
such as degree, betweenness, clustering coefficient, semilocal centrality, 
mean shortest path, page rank, and standard deviation of the shortest path. 
The values for centrality values measures can be added as vertex node 
attributes [calcCentrality] or returned as a R matrix [getCentralityMatrix],
depending on the user’s choice. Any other numerical characteristics, calculated 
for nodes and represented in a matrix form could be stored as node attributes 
as well.

To compare an observed networks vertex centrality values to those of an
equivalently sized  randomised graph, we provide three varying randomisation 
models including G(n,p) Erdos-Renyi model , Barabasi-Albert model, and new 
random graph from a given graph by randomly adding/removing edges
[getRandomGraphCentrality].

Additionally, to allow comparison of network with different structure, we
implemented normalised modularity measure (Parter et al., 2007, 
Takemoto, 2012, Takemoto, 2013, Takemoto and Borjigin, 2011)[normModularity].

The pipeline then proceeds with uncovering the network’s community structure, 
which can be performed via nine different clustering algorithms, simultaneously
[calcAllClustering] or with a chosen algorithm  [calcClustering], community
membership being stored as a vertex attribute. In the situation when the 
network is dense and clusters are large and barely tractable, the reclustering 
can be applied [calcReclusterMatrix]. The obtained community structure can be
visualized with [layoutByCluster] and further tested for robustness 
[getRobustness] by comparing with randomised networks. As a result, consensus
matrix can be estimated [makeConsensusMatrix], which is needed for the next 
step -identifying the "influential" or "bridging" proteins 
(Nepusz et al., 2008).

For this, we enabled a function for estimating "bridgeness" [getBridgeness] 
metric, which takes into account the probability of the vertex to belong to
different communities [getBridgeness], so that the vertices can be ranked under
assumption that more communities the vertex can belong more influence it has 
to the network topology and signaling (Han et al., 2004). "Bridgeness" can be 
plot against any other centrality measure, e.g. semi-local centrality (plot is
implemented), to enable additional view of vertex (protein) importance within 
the network topology.

To get a glance of the molecular signature of a disease or function on the 
network, we implemented the disease-disease overlap analysis by measuring the 
mean shortest distance for each disease (⟨d⟩), using the shortest distance 
between each gene-disease association (GDA) to its next nearest GDA neighbor
(Menche et al., 2015) [calcDiseasePairs], [runPermDisease] to be used for 
obtaining signiificance values. In the presented example of the presynaptic 
network we found a few neurological disorders to overlap with a high 
significance, e.g. AD-PD, SCH-BP, ASD-ID, and, indeed, their comorbidity was
confirmed by the literature. Same analysis can be performed with any in-house
annotation.

To allow the distribution of the specific annotation(s) over clustered graph, 
we enabled overrepresentation analysis, which allows identify the network 
community enriched wit specific function or disease, or any other annotation.

The example illustrates the package functionality for the protein-protein
interaction network for the presynaptic compartment of the synapse generated 
from Synaptic proteome database (Sorokina et al., 2021) with Synaptome.db 
package. The network has 1073 vertices and 3346 edges, the step by step 
analysis shown below.

# Build the network

The package allows building the network from the data frame, where the rows
correspond to the edges of the graph, or for the list of nodes (genes), for 
which the information will be retrieved from the SynaptomeDB.

## Build the network from data frame
The command builds the graph from provided data frame, simplifies the graph
(removing multiple edges and loops) and return its MCC (maximum connected 
component)
```{r network_from_scratch}
file <- system.file("extdata", "PPI_Presynaptic.csv", package = "BioNAR")
tbl <- read.csv(file, sep="\t")
head(tbl)
gg <- buildNetwork(tbl)
summary(gg)

```
## Build network for the node list extracted from SynaptomeDB
The command provides proxy for Synaptome.db package, enabling building the
protein-protein interaction (PPI) network for entire synaptic compartment 
(pre-, or post-synaptic)

```{r net_from_SynaptomeDB}
# Let's get the ID for presynaptic compartment
cid<-match('Presynaptic',getCompartments()$Name) 
cid
# Now we need to collect all the gene IDs for presinaptic  compartment
t<-getAllGenes4Compartment(cid) 
dim(t)
head(t)
# finally, build the graph using respecctive gene EntrezIDs as node IDs
gg<-buildFromSynaptomeByEntrez(t$HumanEntrez) 
summary(gg)

```
## Use the predifined network
Any predefined network stored as a graph file could be loaded for further 
analysis by use of igraph functionality.

```{r net_predefind}
file <- system.file("extdata", "PPI_Presynaptic.gml", package = "BioNAR")
gg1 <- igraph::read.graph(file,format="gml")
summary(gg1)
```

```{r cluster_predefind, include=FALSE}
file <- system.file("extdata", "PPI_cluster.gml", package = "BioNAR")
ggCluster <- igraph::read.graph(file,format="gml")
```



# Annotate the nodes with node attributes
As soon as graph is loaded it can be annotated with any relevant annotations, 
such as protein names, functionality, disease associations, or any customized
annotation (annotate_vertex {BioNAR}. For example, below we use functional
annotation based on two published synaptic studies.

## Gene name
Adding gene names to nodes.
```{r annotate_net}
gg<-annotateGeneNames(gg)
summary(gg)
head(V(gg))
head(V(gg)$GeneName)
```

## Diseases

Adding diseases associations to genes linked to Human Disease Ontology 
(HDO) terms extracted from the package
(topOnto.HDO.db)[https://github.com/hxin/topOnto.HDO.db].
```{r annotate_topOnto}
afile<-system.file("extdata", "flatfile_human_gene2HDO.csv", package = "BioNAR")
dis <- read.table(afile,sep="\t",skip=1,header=FALSE,strip.white=TRUE,quote="")
gg <- annotateTopOntoOVG(gg, dis)
summary(gg)

```

## Schizopherina related synaptic gene functional annotation.

Adding the annottation curated from 
Lips et al., (2012) <doi:10.1038/mp.2011.117>.

```{r annotate_Shanno}
sfile<-system.file("extdata", "SCH_flatfile.csv", package = "BioNAR")
shan<- read.table(sfile,sep="\t",skip=1,header=FALSE,strip.white=TRUE,quote="")
sgg<-annotateSCHanno(gg,shan)
summary(sgg)
```

## Presynaptic functional annotation

Adding the presynaptic genes functional annotation derived from 
Boyken at al. (2013) <doi:10.1016/j.neuron.2013.02.027>.

```{r annotate_Chua, eval=FALSE}
sfile<-system.file("extdata", "PresynAn.csv", package = "BioNAR")
pres <- read.csv(sfile,skip=1,header=FALSE,strip.white=TRUE,quote="")
sgg <- annotatePresynaptic(gg, pres)
summary(sgg)
```

## GO

Adding functionality from GO: BP, MF,CC
```{r annotate_go}
sfile<-system.file("extdata", "flatfile.go.BP.csv", package = "BioNAR")
goBP <- read.table(sfile,sep="\t",skip=1,header=FALSE,strip.white=TRUE,quote="")
sgg <- annotateGoBP(gg, goBP)
summary(sgg)

sfile<-system.file("extdata", "flatfile.go.MF.csv", package = "BioNAR")
goMF <- read.table(sfile,sep="\t",skip=1,header=FALSE,strip.white=TRUE,quote="")
sgg <- annotateGoMF(gg, goMF)
summary(sgg)

sfile<-system.file("extdata", "flatfile.go.CC.csv", package = "BioNAR")
goCC <- read.table(sfile,sep="\t",skip=1,header=FALSE,strip.white=TRUE,quote="")
sgg <- annotateGoCC(gg, goCC)
summary(sgg)
```

# Esimate node centrality measures

## Estimate centrality measures with values added as node attributes. 

Centrality measures are as following:
* DEG - degree, 
* BET - betweenness, 
* CC - clustering coefficient, 
* SL - semilocal centrality, 
* mnSP - mean shortest path, 
* PR - page rank, 
* sdSP - standard deviation of the shortest path.

```{r graph_cent}
gg <- calcCentrality(gg)
summary(gg)
```

## Get node centralities as a matrix. 

In the case you don't want save centrality values on the graph, for example 
you going to provide different names for the attributes, the same centrality
measures as before could be obtained in a matrix form:

```{r matrix_cent}
mc <- getCentralityMatrix(gg)
head(mc)
```

## Get the centrality measures for random graph

Sometimes one needs to compare the graph properties of the given network with the ones of the randomized graph. Command below provides three ways of randomization, including G(n,p) Erdos-Renyi model , Barabasi-Albert model and new random graph from a given graph by randomly adding/removing edges.

```
{r}
ggrm <- getRandomGraphCentrality(sgg, type = c("cgnp"))
head(ggrm)
```

## Power law fit

To examine the networks for underlying structure (i.e. not random), one can 
test network's degree distribution for evidence of scale-free structure and 
compare it against the randomised network model. For this we used the R 
“PoweRlaw” package (version 0.50.0) (Gillespie, 2015).
In the example below we found evidence for disassortative mixing 
(Newman, 2002), i.e. a preference for high-degree genes to attach to 
low-degree gene(presynaptic: -0.16).

```{r powerLaw,fig.height=8,fig.width=8,dpi=56}
pFit <- fitDegree( as.vector(igraph::degree(graph=gg)),threads=1,Nsim=5,
                    plot=TRUE,WIDTH=2480, HEIGHT=2480)
```

## Get entropy rate

Evidence for scale-free structure can also be tested by performing a 
perturbation analysis on each network. In this analysis each protein is 
being perturbed through over-expression (red) and under-expression (green), 
with the global entropy rate (SR) after each proteins pertubation being 
plotted against the log of the proteins degree, as shown at the plot below. 
In the example network we observe a bi-modal response, between gene 
over-expression and degree, and opposing bi-phasic response relative to
over/under-expression between global entropy rate and degree. This type of
bi-modal, bi-phasic behaviour has been observed only in networks with 
scale-free or approximate scale-free topology (Teschendorff et al, 2014). 

```{r ent_rate,fig.height=8,fig.width=8,dpi=56}
ent <- getEntropyRate(gg)
ent
SRprime <- getEntropy(gg, maxSr = NULL)
head(SRprime)
plotEntropy(SRprime, subTIT = "Entropy", SRo = ent$SRo, maxSr = ent$maxSr)
```

## Get modularity. Normalised modularity.
Normalised modularity allows comparison the networks with different structure.
Qm based on the previous studies by Parter et al., 2007, Takemoto, 2012,
Takemoto, 2013, Takemoto and Borjigin, 2011, which was defined as:

$$Qm = \frac{Q_{real}-Q_{rand}}{Q_{max}-Q_{rand}}$$

Where $Q_{real}$ is the network modularity of a real-world signaling network 
and, $Q_{rand}$ is the average network modularity value obtained from 
10,000 randomized networks constructed from its real-world network. 
$Q_{max}$ was estimated as: $$Q_{max}=1 − \frac{1}{M}$$, where $M$ is the 
number of modules in the real network.

Randomized networks were generated from a real-world network using the
edge-rewiring algorithm (Maslov and Sneppen, 2002).
```{r norm_mod}
nm<-normModularity(gg,alg='louvain')
nm
```

# Clustering
Clustering, or community detection, in networks has been well studied in the 
field of statistical physics with particular attention to methods developed 
for social science networks. The underlying assumption(s) of what makes a 
community in social science, translates remarkably well to what we think of 
as a community (sub-complex, module or cluster) in PPI networks
The possible algorithms of choice implemented in the package are:
* "lec"('Leading-Eigenvector, Newman, 2006), 
* "wt"(Walktrap, Pons & Latapy, 2006), 
* "fc"(Fast-Greedy Community' algorithm, Clauset et al., 2004), 
* "infomap" (InfoMAP, Rosvall et al., 2007; Rosvall et al., 2010), 
* "louvain" (Louvain, Blondel et al., 2008), 
* "sgG1", "sgG2", "sgG5"(SpinGlass, Reichardt & Bornholdt). 
For each algorithm of interest the membership could be obtained with 
`calcMembership` command.
All algorithm implementations, apart from Spectral were performed using the 
publicly available R package igraph (Csardi & Nepusz, 2006) (R version 3.4.2, 
igraph version 1.1.2). Parameters used in the fc, lec, sg, wt and lourvain 
algorithms were chosen as to maximise the measure Modularity 
(Newman & Girvan, 2004); infomap seeks the optimal community structure in 
the data by maximising the objective function called the Minimum Description 
Length (Rissanen, 1978; Grwald et al., 2005)

```{r cluster.mem}
# choose one algorithm from the list
alg = "louvain"
mem <- calcMembership(gg, alg)
pander(head(mem))
```

Due to internal random initialisation consequtive invocation of the same 
algorithm could produce slightly different community structures:
```{r cluster.mem2}
mem2 <- calcMembership(gg, alg)
idx<-match(mem$names,mem2$names)
idnx<-which(mem$membership!=mem2$membership[idx])
pander(head(cbind(mem[idnx,],mem2[idx[idnx],])))
```

To avoid inconsistency if downstream analysis we provide two additional 
functions `calcClustering` and `calcAllClustering` that use calcMembership to 
calculate community memberships and store them within the graph vertices 
attributes named after the algorithm. They also calculate modularity values 
and store them on graph attributes named after the algorithme. The difference 
between `calcClustering` and `calcAllClustering` is that `calcAllClustering` 
allows to calculate memberships for all clustering algorithms simultaneously 
(may take time), and store them on the graph vertices, while `calcClustering`  
command will work for specific algorithm. 

```{r cluster}
gg <- calcClustering(gg, alg)
summary(gg)
```

Membership data could be obtained from the graph vertex attribute:

```{r get.attr}
mem.df<-data.frame(names=V(gg)$name,membership=as.numeric(V(gg)$louvain))
```

To compare different clustering algorithms summary matrix can be calculated 
with the follownig values:
1. maximum Modularity obtained (mod), 
2. number of detected communities (C), 
3. the number of communities with size (Cn1) equal to 1, 
4. the number of communities >= 100 (Cn100), 
5. the fraction of edges lying between communities (mu), 
6. the size of the smallest community (Min. C), 
7. the size of the largest community (Max. C), 
8. the average ( Mean C), median (Median C), 
9. first quartile (1st Qu. C), and 
10. third quartile (3rd Qu. C) of the community size. 

```{r calcAllClustering,eval=FALSE}
ggc <- calcAllClustering(gg)
```
```{r calcAllClustering.hid,eval=TRUE,include=FALSE}
ggc <- ggCluster
```
```{r clusterSummary,eval=TRUE}
m<-clusteringSummary(ggc,att=c('lec','wt','fc',
                                'infomap','louvain',
                                'sgG1','sgG2','sgG5'))
pander(m)
```

It is often useful to be able to visualize clusters on the graph. The simples 
way is to color each cluster in its own color and plot the graph:
```{r plot.color.graph,fig.height=8,fig.width=8,dpi=56}
palette <- distinctColorPalette(max(as.numeric(mem.df$membership)))
plot(gg,vertex.size=3,layout=layout_nicely,
        vertex.label=NA,
        vertex.color=palette[as.numeric(mem.df$membership)],
        edge.color='grey95')
legend('topright',legend=names(table(mem.df$membership)),
        col=palette,pch=19,ncol = 2)
```

On the plot we can see some distinctive clusters but most nodes are 
indistinguishable within the central part of the plot. So we could layout 
graph clusterwise:
```{r plot.clusterwise.graph,fig.height=8,fig.width=8,dpi=56}
lay<-layoutByCluster(gg,mem.df,layout = layout_nicely)
plot(gg,vertex.size=3,layout=lay,
        vertex.label=NA,
        vertex.color=palette[as.numeric(mem.df$membership)],
        edge.color='grey95')
legend('topright',legend=names(table(mem.df$membership)),
        col=palette,pch=19,ncol = 2)
```

It is also possible to visualize the interaction between communities:
```{r ploc.cluster.communities,fig.height=8,fig.width=8,dpi=56}
idx<-match(V(gg)$name,mem.df$names)
cgg<-getCommunityGraph(gg,mem.df$membership[idx])
D0 = unname(degree(cgg))
plot(cgg, vertex.size=sqrt(V(cgg)$size), vertex.cex = 0.8,
vertex.color=round(log(D0))+1,layout=layout_with_kk,margin=0)
```

### Reclustering
Reclustering a clustered graph using the same, or different, 
clustering algorithm:
```{r recluster}
remem<-calcReclusterMatrix(gg,mem.df,alg,10)
head(remem)
```

And we can apply second order clustering layout:
```{r plot.recluster.layout,fig.height=8,fig.width=8,dpi=56}
lay<-layoutByRecluster(gg,remem,layout_nicely)
plot(gg,vertex.size=3,layout=lay,
        vertex.label=NA,
        vertex.color=palette[as.numeric(mem.df$membership)],
        edge.color='grey95')
legend('topright',legend=names(table(mem.df$membership)),
        col=palette,pch=19,ncol = 2)
```


## Consensus matrix
For assessing the robustness of the clustering randomization study can be 
performed, which applies the same clustering algorithm to N perturbed networks 
and returns the consensus matrix where each pair of nodes will be assigned 
the frequency to belong to the same cluster. 

Where 'alg' gives the name of the clustering algorithm and 'type' the 
sampling scheme (1 sample edges, and 2 sample nodes) to use, 'mask' the 
percentage of edges or nodes to be mask, 'reclust' whether reclustering 
should be performed on the community set found, 'Cnmin' minimum cluster 
size and 'Cnmax' the maximum cluster size above which reclustering will be 
preformed (if reClust=TRUE). 
```{r cons_mat}
#Build consensus matrix for louvain clustering
conmat <- makeConsensusMatrix(gg, N=5,
                                alg = alg, type = 2, 
                                mask = 10,reclust = FALSE, 
                                Cnmax = 10)

```
For the sake of timing we use only five 
randomisatio round (\code{N=5}), for the real analysis you should use at least
\code{N=100}.

##Consensus matrix value distribution

Consensus matrix value distribution could be visualised in the following way:
```{r plot.conmat.ecdf,fig.height=8,fig.width=8,dpi=56}
steps <- 100
Fn  <- ecdf(conmat[lower.tri(conmat)])
X<-seq(0,1,length.out=steps+1)
cdf<-Fn(X)
dt<-data.frame(cons=X,cdf=cdf)
ggplot(dt,aes(x=cons,y=cdf))+geom_line()+
        theme(            
        axis.title.x=element_text(face="bold",size=rel(2.5)),
        axis.title.y=element_text(face="bold",size=rel(2.5)),
        legend.title=element_text(face="bold",size=rel(1.5)),
        legend.text=element_text(face="bold",size=rel(1.5)),
        legend.key=element_blank())+
    theme(panel.grid.major = element_line(colour="grey40",size=0.2),
            panel.grid.minor = element_line(colour="grey40",size=0.1),
            panel.background = element_rect(fill="white"),
            panel.border = element_rect(linetype="solid",fill=NA))
```


## Cluster robustness

Cluster robustness assess the robustness of the obtained clusters and 
could help to evaluate the "goodness" of chosen clustering algorithm.
```{r clcons}
clrob<-getRobustness(gg, alg = alg, conmat)
pander(clrob)
```


# Bridgeness 

Bridging proteins are known to interact with many neighbours simultaneously,
organise function inside communities they belong to, but also affect/influence
other communities in the network (Nepusz et al., 2008).Bridgeness is 
estimated from the consensus clustering estimated above and vertex degree 
to calculate the vertex’s community membership, i.e. the probability of 
the specific node to belong to every communities obtained by given 
clustering algorithm.

The Bridgeness measure lies between 0 - implying a vertex clearly belongs in 
a single community, and 1 - implying a vertex forms a 'global bridge' across 
every community with the same strength. 

```{r calc.bridge}
br<-getBridgeness(gg,alg = alg,conmat)
pander(head(br))
```

## Bridgeness plot

Semi-local centrality measure (Chen et al., 2011) also lies between 0 and 
1 indicating whether protein is important globally or locally. By plotting
Bridgeness against semi-local centrality we can categorises the influence 
each protein found in our network has on the overall network structure:
* Region 1, proteins having a 'global' rather than 'local' influence in the 
network (also been called bottle-neck bridges, connector or 
kinless hubs (0<Sl<0.5; 0.5<Br<1). 
* Region 2, proteins having 'global' and 'local' influence 
(0.5<Sl<1, 0.5<Br<1). 
* Region 3, proteins centred within the community they belong to, 
but also communicating with a few other specific communities 
(0<Sl<0.5; 0.1<Br<0.5). 
* Region 4, proteins with 'local' impact , primarily within one or 
two communities (local or party hubs, 0.5<Sl<1, 0<Br<0.5).
```{r plot.bridgeness,fig.height=8,fig.width=8,dpi=56}
scale <- function(x, VALUE=NULL){
    
    x = as.numeric(as.vector(x))
    
    xmin <- min(x,na.rm=TRUE)
    xmax <- max(x,na.rm=TRUE)
    
    if( is.null(VALUE) ){
    
    x  <- x-xmin
    x  <- ifelse(!is.na(x), x/(xmax-xmin), NA) 
    
    return(x)
    }
    
    value = as.numeric(as.vector(value)[1])
    value = value-xmin
    value = ifelse(!is.na(value), value/(xmax-xmin), NA) 
    return(value)
}
VIPs=c('8495','22999','8927','8573','26059','8497','27445','8499')
# VIPs=c('81876','10890','51552','5874','5862','11021','54734','5865','5864',
#        '9522','192683','10067','10396','9296','527','9114','537','535',
#        '528','51382','534','51606','523','80331','114569','127262','57084',
#        '57030','388662','6853','6854','8224','9900','9899','9145','9143',
#        '6855','132204','6857','127833','6861','529','526','140679','7781',
#        '81615','6844','6843')
indx   <- match(V(gg)$name,VIPs)
group <- ifelse( is.na(indx), 0,1)
MainDivSize <- 0.8
xmin        <- 0
xmax        <- 1
ymin        <- 0
ymax        <- 1
Xlab <- "Semilocal Centrality (SL)" 
Ylab <- "Bridgeness (B)"
X    <- as.numeric(igraph::get.vertex.attribute(gg,"SL",V(gg)))
X    <- scale(X)
Y       <- as.numeric(as.vector(br[,dim(br)[2]])) 
lbls <- ifelse(!is.na(indx),V(gg)$GeneName,"")
dt<-data.frame(X=X,Y=Y,vips=group,entres=V(gg)$name,name=V(gg)$GeneName)
dt_vips<-dt[dt$vips==1,]
dt_res<-dt[dt$vips==0,]
##--- baseColor of genes
baseColor="royalblue2"

##--- SPcolor, colour highlighting any 'specical' genes
SPColor="royalblue2"

##--- PSDColor, colour of core PSD95 interactor genes
PSDColor="magenta"

ggplot(dt,aes(x=X,y=Y,label=name))+#geom_point()+
    geom_point(data=dt_vips,
                aes(x=X,y=Y),colour=baseColor,size = 7,
                shape=15,show.legend=FALSE)+
    geom_point(data=dt_res,
                aes(x=X,y=Y, alpha=(X*Y)), size = 3,
                shape=16,show.legend=FALSE)+
    geom_label_repel(aes(label=as.vector(lbls)),
                        fontface='bold',color='black',
                        fill='white',box.padding=0.1,
                        point.padding=NA,label.padding=0.15,
                        segment.color='black',
                        force=1,size=rel(3.8),show.legend=FALSE,
                        max.overlaps=200)+
        labs(x=Xlab,y=Ylab,title=sprintf("%s",alg))+
    scale_x_continuous(expand = c(0, 0), limits = c(xmin, xmax)) + 
    scale_y_continuous(expand = c(0, 0), limits = c(ymin, ymax))+
    theme(            
        axis.title.x=element_text(face="bold",size=rel(2.5)),
        axis.title.y=element_text(face="bold",size=rel(2.5)),
        legend.title=element_text(face="bold",size=rel(1.5)),
        legend.text=element_text(face="bold",size=rel(1.5)),
        legend.key=element_blank())+
    theme(panel.grid.major = element_line(colour="grey40",size=0.2),
            panel.grid.minor = element_line(colour="grey40",size=0.1),
            panel.background = element_rect(fill="white"),
            panel.border = element_rect(linetype="solid",fill=NA))+
        geom_vline(xintercept=0.5,colour="grey40",size=MainDivSize,
                    linetype=2,show.legend=FALSE)+
    geom_hline(yintercept=0.5,colour="grey40",size=MainDivSize,
                linetype=2,show.legend=FALSE)

```

##Interactive view of bridgeness plot

With the help from \code{\link[plotly]{ggplotly}} function we could make the
plot more interactive.
```{r plotly,fig.height=6,fig.width=6}
library(plotly)
p<-ggplot(dt,aes(x=X,y=Y,label=name))+#geom_point()+
    geom_point(data=dt_vips,
                aes(x=X,y=Y),colour=baseColor,shape=15,show.legend=FALSE)+
    geom_point(data=dt_res,
                aes(x=X,y=Y, alpha=(X*Y)),shape=16,show.legend=FALSE)+
        labs(x=Xlab,y=Ylab,title=sprintf("%s",alg))+
    scale_x_continuous(expand = c(0, 0), limits = c(xmin, xmax)) + 
    scale_y_continuous(expand = c(0, 0), limits = c(ymin, ymax))+
    theme(            
        axis.title.x=element_text(face="bold",size=rel(2.5)),
        axis.title.y=element_text(face="bold",size=rel(2.5)),
        legend.title=element_text(face="bold",size=rel(1.5)),
        legend.text=element_text(face="bold",size=rel(1.5)),
        legend.key=element_blank())+
    theme(panel.grid.major = element_line(colour="grey40",size=0.2),
            panel.grid.minor = element_line(colour="grey40",size=0.1),
            panel.background = element_rect(fill="white"),
            panel.border = element_rect(linetype="solid",fill=NA))+
        geom_vline(xintercept=0.5,colour="grey40",size=MainDivSize,
                    linetype=2,show.legend=FALSE)+
    geom_hline(yintercept=0.5,colour="grey40",size=MainDivSize,
                linetype=2,show.legend=FALSE)
ggplotly(p)
```

# Disease/annotation pairs

Given that Disease associated genes are connected within the graph, the 
common question is to check whether the networks for two different 
diseases are overlapping, which may indicate the common molecular mechanisms. 
Same is valid for any pair of annotations, e.g. one would ask if two 
different biological functions are related.

To address the question, we have utilised the algorithm from Menche et al, 
which estimates the minimal shortest paths between two distinct annotations 
and compare it to the randomly annotated graph.

Below example shows the estimation of disease separation for the 
following diseases: DOID:10652 (Alzheimer's_disease),  (bipolar_disorder),
DOID:12849 (autistic_disorder), DOID:1826 (epilepsy)
Command `calcDiseasePairs` quickly estimates the two annotation separation on 
the graph and compares it with one randomly reannotated graph. This could be 
used for initial guess of the relationships between the annotations.

To assess the significance of the obtained separation values the command
`runPermDisease` should be used, where the user can dfine the number of 
permutations. The command execution will take a while, depending on this 
number and return the table with p-value, p.adjusted, q-value and 
Bonferroni test.
```{r disPairs,warning=FALSE,message=FALSE}
p <- calcDiseasePairs(
    gg,
    name = "TopOnto_OVG_HDO_ID",
    diseases = c("DOID:10652","DOID:3312"),
    permute = "r"
)
pander(p$disease_separation)

r <- runPermDisease(
    gg,
    name = "TopOnto_OVG_HDO_ID",
    diseases = c("DOID:10652","DOID:3312"),
    Nperm = 100,
    alpha = c(0.05, 0.01, 0.001)
)

pander(r$Disease_overlap_sig)
```

# Session Info
```{r sessionInfo, echo=FALSE, results='asis', class='text', warning=FALSE}
library(devtools)
si<-devtools::session_info()
cat('Platform\n\n')
pander::pander(si$platform)
cat('Packages\n\n')
knitr::kable(as.data.frame(si$packages)[,c('ondiskversion',
                                            'loadedversion','date',
                                            'source')],align = c('l','l'))
```

