---
title: "Laplacian Renormalization Group (LRG) Analysis"
author:
- name: "Colin Mclean"
- name: "Anatoly Sorokin"
- name: "T. I. Simpson"
- name: "J. Douglas Armstrong"
- name: "Oksana Sorokina"
date: "Edited: 2025; Compiled: `r format(Sys.time(), '%B %d, %Y')`"
package: BioNAR
vignette: >
  %\VignetteIndexEntry{Laplacian Renormalization Group (LRG) Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    number_sections: yes
    toc: yes
    toc_depth: 4
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
library(knitr)
library(BioNAR)
library(ggplot2)
library(pander)
library(ggrepel)
library(randomcoloR)
library(clustAnalytics)
```
```{r functions, include=FALSE,cache=FALSE}
alORA<-paste0('|p{0.01\\textwidth}',
           '|r',                #cl
           '|r',                #FL
           '|r',                #N
           '|r',                #Fn
           '|r',                #Cn
           '|r',                #Mu
           '|r',                #OR
           '|r',                #Fe
           '|r',                #Fc
           '|r',                #pval
           '|r',                #padj
           '|r',                #palt
           '|r|'                #paltadj
)

```


# Introduction

# Presynaptic PPI network analysis
## Loading network
```{r ppi.read.data}
file <- system.file("extdata", "PPI_Presynaptic.gml", package = "BioNAR")
gg <- igraph::read_graph(file,format="gml")
gg<-calcAllClustering(gg)
summary(gg)
```

```{r ppi.powerlaw.plot}
pFit <- fitDegree( as.vector(igraph::degree(graph=gg)),threads=1,Nsim=5,
plot=TRUE,WIDTH=2480, HEIGHT=2480)
```

Calculation of the optimal coarse-grained graph could be time-consuming ()
```{r ppi.run.LRG}
res<-runLRG(gg,cooling_rate=0.85,seed=100,
            complex=FALSE, method="eigen", max_iter=100, dt=1,
            comb='concat',simplify=FALSE)
```

```{r ppi.LRG.finplot}
plot(res$finplot)
```

```{r ppi.LRG.tpsplot}
plot(res$tps$pt)
```


```{r ppi.LRG.logplot}
plot(res$logplot)
```

Supernode size distribution:
```{r ppi.LRG.supernode.stats}
kable(table(res$R$supernodes))
```

Supernode composition:
```{r ppi.LRG.supernode.comp}
kable(V(res$R$gg)$GeneName[which(res$R$supernodes>1)])
```

Degree distribution tale comparison, original graph:
```{r ppi.LRG.degree.tails.orig}
kable(sort(degree(res$gg),decreasing=TRUE)[1:15])
```

coarse-grained graph:
```{r ppi.LRG.degree.tails.cg}
kable(sort(degree(res$R$gg),decreasing=TRUE)[1:15])
```

## Cluster analysis
```{r ppi.LRG.cSummary}
gg<-applpMatrixToGraph(gg,res$R$mapping)
csum<-clusteringSummary(gg,att = c("lec", "wt", "fc", "infomap", "louvain", "sgG1", "sgG2", "sgG5", "spectral",'LRG'))
lcORA<-list()
kable(csum)
```


### Walktrap community
```{r ppi.LRG.wt.ora}
clORA<-clusterORA(gg,'wt','LRG')
lcORA[['wt']]<-clORA[!is.infinite(clORA$OR),]
```

### Spin-glass model community
```{r ppi.LRG.sgG1.ora}
clORA<-clusterORA(gg,'sgG1','LRG')
lcORA[['sgG1']]<-clORA[!is.infinite(clORA$OR),]
```

```{r ppi.LRG.sgG2.ora}
clORA<-clusterORA(gg,'sgG2','LRG')
lcORA[['sgG2']]<-clORA[!is.infinite(clORA$OR),]
```

```{r ppi.LRG.sgG5.ora}
clORA<-clusterORA(gg,'sgG5','LRG')
lcORA[['sgG5']]<-clORA[!is.infinite(clORA$OR),]
```

### Infomap community
```{r ppi.LRG.infomap.ora}
clORA<-clusterORA(gg,'infomap','LRG')
lcORA[['infomap']]<-clORA[!is.infinite(clORA$OR),]
```

### Louvain
```{r ppi.LRG.louvain.ora}
clORA<-clusterORA(gg,'louvain','LRG')
lcORA[['louvain']]<-clORA[!is.infinite(clORA$OR),]
```

### Spectral
```{r ppi.LRG.spectral.ora}
clORA<-clusterORA(gg,'spectral','LRG')
lcORA[['spectral']]<-clORA[!is.infinite(clORA$OR),]
```

```{r ppi.LRG.ora.table}
tORA<-do.call(rbind,lcORA)
kable(tORA[,-c(1,9,10,17)])
```

# Diseasesome analysis

## Prepare network
```{r disease.prep.net}
rm(res)
gd<-findLCC(diseasome)
gd<-calcAllClustering(gd)
summary(gd)
```

```{r disease.powerlaw.plot}
pFit <- fitDegree( as.vector(igraph::degree(graph=gd)),threads=1,Nsim=5,
plot=TRUE,WIDTH=2480, HEIGHT=2480)
```


Calculation of the optimal coarse-grained graph could be time-consuming ()
```{r disease.run.LRG}
res<-runLRG(gd,cooling_rate=0.85,seed=100,
            complex=FALSE, method="eigen", max_iter=100, dt=1,
            comb='concat',simplify=FALSE)
```

```{r disease.LRG.finplot}
plot(res$finplot)
```

```{r disease.LRG.tpsplot}
plot(res$tps$pt)
```

```{r disease.LRG.logplot}
plot(res$logplot)
```

Supernode size distribution:
```{r disease.LRG.supernode.stats}
kable(table(res$R$supernodes))
```

Supernode composition:
```{r disease.LRG.supernode.comp}
kable(V(res$R$gg)$GeneName[which(res$R$supernodes>1)])
```

Degree distribution tale comparison, original graph:
```{r disease.LRG.degree.tails.orig}
kable(sort(degree(res$gg),decreasing=TRUE)[1:15])
```

coarse-grained graph:
```{r disease.LRG.degree.tails.cg}
kable(sort(degree(res$R$gg),decreasing=TRUE)[1:15])
```

## Cluster analysis
```{r disease.LRG.cSummary}
gd<-applpMatrixToGraph(gd,res$R$mapping)
csum<-clusteringSummary(gd,att = c("lec", "wt", "fc", "infomap", "louvain", "sgG1", "sgG2", "sgG5", "spectral",'LRG'))
lcORA<-list()
kable(csum)
```


### Walktrap community
```{r disease.LRG.wt.ora}
clORA<-clusterORA(gd,'wt','LRG')
lcORA[['wt']]<-clORA[!is.infinite(clORA$OR),]
```

### Spin-glass model community
```{r disease.LRG.sgG1.ora}
clORA<-clusterORA(gd,'sgG1','LRG')
lcORA[['sgG1']]<-clORA[!is.infinite(clORA$OR),]
```

```{r disease.LRG.sgG2.ora}
clORA<-clusterORA(gd,'sgG2','LRG')
lcORA[['sgG2']]<-clORA[!is.infinite(clORA$OR),]
```

```{r disease.LRG.sgG5.ora}
clORA<-clusterORA(gd,'sgG5','LRG')
lcORA[['sgG5']]<-clORA[!is.infinite(clORA$OR),]
```

### Infomap community
```{r disease.LRG.infomap.ora}
clORA<-clusterORA(gd,'infomap','LRG')
lcORA[['infomap']]<-clORA[!is.infinite(clORA$OR),]
```

### Spectral
```{r disease.LRG.spectral.ora}
clORA<-clusterORA(gd,'spectral','LRG')
lcORA[['spectral']]<-clORA[!is.infinite(clORA$OR),]
```

```{r disease.LRG.ora.table}
tORA<-do.call(rbind,lcORA)
kable(tORA[,-c(1,9,10,17)])
```


